---
title: 'Activity: Popular Names'
author: "Abdulaziz Bokijonov"
date: "5 June 2023"
output:
  html_notebook: default
  html_document:
    df_print: paged
  pdf_document: default
---

# Grading

Assignment is worth a total of 10 points

-   [1 point] Turn in BOTH RNotebook (.nb.html) and the RMarkdown (.Rmd) file to Canvas
-   [1 point] Changes pushed to GitHub
-   [1 point] Step 1: show how you used summary functions to explore the raw data set
-   [1 point] Step 3 and 4: attempt to answer all questions about the data
-   [1 point] Step 5: explanation of data wrangling steps in English (not code)
-   [3 points] Step 7: Data wrangling
    -   correct use of `filter()`
    -   correct use of `group_by()`
    -   correct use of `summarise()`
-   [2 points] `ggplot()` graph of first names

# Set Up

```{r, message = F, warning = F}
# load necessary packages
library(tidyverse)
library(mosaic)
library(esquisse)


# Load the BabyNames data into RStudio environment (from dcData package)
data("BabyNames", package = "dcData")
```

# Activity

**Goal**: By the end of the activity, each team will create a graph (like Figure A.1 on p 159) showing ups and downs in the popularity of the names at your table according to the U.S. social security registry.

For steps 5-7, pick any FIVE names you wish. I choose Arnaldo, Ahmid, Nicho, Jennaveve, and Christy. You may choose these names or any others that you would like.

### Step 1

*Examine the data you have at hand - for this project the data are the ones available in the `BabyNames` data set. Use at least 2 different functions to examine your data. Then answer the questions: What are the variables are available? What is a case?*

Cases are the babies whose names are being studies. The variables available are their names, sex, year and the number of babies with that name

```{r}
## Look at the data a few different ways (see Ch 9 for more).
head(BabyNames)
tail(BabyNames)
summary(BabyNames)
```

### Step 3

##### What variable(s) from the raw data table do not appear at all in the graph?

A variable sex does not appear on the graph

##### What variable(s) in the graph are similar to corresponding variables in the raw data table, but might have been transformed in some way?

Popularity variable on the graph is the same as the count variable on the table.

### Step 4

##### Have cases been filtered out?

Only 5 names/cases from the dataset were chosen based on the personal preference.

##### Have cases been grouped and summarized within groups in any way?

Cases have been grouped by the year.

##### Have any new variables been introduced? If so, what the relationship between the new variables and existing variables?

No, but like mentioned above, popularity variable was introduced which substitutes count variable on the table.

### Step 5

Using the BabyNames data set, filter the data to display just the 5 names that you picked. Then group the cases by year and name, and count how many total births per year were recorded for each of the names (regardless of sex).

I will use group_by function to group the data by name and will use filter function to select 5 names.

### Step 7

```{r}
## Create my glyph ready data set here
GlyphReadyForm <- BabyNames %>%
  group_by(year, name) %>%
  filter(
    name == "Sarah" |
    name == "Rachael" |
    name == "Adam" |
    name == "Daisy" |
    name == "Ana"
    )

# inspect the result
GlyphReadyForm
```

### Graph It!

Here's the sample code provided by the book:

```         
GlyphReadyForm %>%
  ggplot(aes(x = year, y = total, group = name)) + 
  geom_line(size = 1, alpha = 0.5, aes(color = name)) + 
  ylab("Popularity") + 
  xlab("Year") +
  ggtitle("Popularity of Names Over Time")
```

If you prepared your `GlyphReadyForm` data set correctly, you should just be able to copy and paste this code. Alternatively, you can use `mplot` or `esquisse` to create your graph. If you choose to use these interactive functions remember, do not include the `mplot` or `esquisse` functions in your RMarkdown document. Instead, use these functions in the console and copy the generated code into your RMarkkdown document. In addition, we want lines, not dots. See examples below.

```{r}
#My code here 
GlyphReadyForm %>%
  ggplot(aes(x = year, y = count, group = name)) + 
  geom_line(linewidth = 1, alpha = 0.5, aes(color = name)) + 
  ylab("Popularity") + 
  xlab("Year") +
  ggtitle("Popularity of Names Over Time")

```

